---
title: "#30DayChartChallenge day24: Theme Day UN Woman"
author: "Federica Gazzelloni"
format: html
editor: visual
execute: 
  warning: false
  message: false
---


> Gazzelloni F. (2023), Data Visualization: UN Woman

Data for this visualization is from the [un.org](https://population.un.org/), to download the data: [https://population.un.org/wpp/Download/Standard/Fertility/](https://population.un.org/wpp/Download/Standard/Fertility/).

The theme for the challenge is **Theme Day: UN Woman** and I chosen a package made by **Cédric Vidonne**, which provides a ggplot2 theme and a set of colour palettes for making charts and graphics based on **UNHCR Data Visualization Guidelines**.

To install the package:
```{r}
install.packages("remotes")
remotes::install_github("vidonne/unhcrthemes")
```


Load necessary packages:
```{r}
library(tidyverse)
library(scales)
library(unhcrthemes)
```


Download and load the data:
```{r}
library(readxl)
df <- read_excel("data/WPP2022_FERT_F04_BIRTHS_BY_5-YEAR_AGE_GROUPS_OF_MOTHER.xlsx", 
    sheet = "Estimates", 
    skip = 16)
df%>%names
```


```{r}
df1 <- df%>%
  janitor::clean_names()%>%
  pivot_longer(cols = 13:20)%>%
  select(region=region_subregion_country_or_area,
         year,name,value)%>%
  mutate(value2=as.numeric(value),
         value2=round(value2),
         name=gsub("x","",name),
         name=gsub("_","-",name))%>%
  rename(age=name) %>%
  filter(!is.na(year))

df1%>%head;
df1%>%dim
```

164160      4
```{r}
df1%>%summary()
```


```{r}
options(scipen=999)
df1%>%
  filter(region=="WORLD",
         year==2021) %>%
  ggplot(aes(x = age, y = value2)) +
  geom_bar(stat = "identity", fill = "#0072C6") +
  labs(
    title = "World: Number of births classified by five-year age groups of mother",
    subtitle="Data are presented in thousands (K)",
    x = "Age category",
    y = "Value",
    caption="DataSource:Fertility standard projections (population.un.org)\n#30DayChartChallenge day24: Theme Day: UN Woman | DataViz: Federica Gazzelloni"
  ) +
  scale_y_continuous(labels = label_number(suffix = "K", scale = 1e-3))+
  unhcrthemes::theme_unhcr()
```

```{r}
library(showtext)
showtext.auto(enable = FALSE)
ggsave("day_14_UNWoman.png",
       bg="white",
       width = 7,height = 5)
```

